---
title: 'IMT 573: PSet 7 - Logistic Regression and Prediction'
author: "Iu-Yu Melody Chang"
output: html_document
---

# Instructions

Before beginning this assignment, please ensure you have access to R
and RStudio.

1. Update the PSet number and title in the `title:` field.

1. Replace the "Insert Your Name Here" text in the `author:` field
   with your own full name. 

1. Be sure to include well-documented (e.g. commented) code chucks,
   figures and clearly written text-chunk explanations as
   necessary. Any figures should be clearly labeled and appropriately
   referenced within the text.

1. When you have completed the assignment and have **checked** that
   your code both runs in the Console and knits correctly when you
   click `Knit to HTML`, rename the R Markdown file to
   `YourLastName-YourFirstName-psetX.Rmd`, where `X` is the PSet number, 
   knit the .rmd file as an html, and submit both the .rmd and the .html files 
   on Canvas.
   
1. List any collaborators in the section below.

# Collaborators
- Bella, Yuki discussed with me!

# Setup

Do whatever setup you do here, such as loading libraries

```{r setup, message=FALSE}
library("tidyverse")
library("margins")
```

# Question 1

## Question 1.1.1
Read the Titanic file and check the NAs in the file.
```{r}
titanic <- read.csv("titanic.csv.bz2")

```
## Question 1.1.2
Only age has 263 missing values. In survived, pclass, sex variable, there aren't any missing values.There are 263 NAs in age variable.
```{r}
titanic <- read.csv("titanic.csv.bz2")
summary(titanic)
sum(is.na(titanic$survived))
sum(is.na(titanic$pclass))
sum(is.na(titanic$sex))
sum(is.na(titanic$age))
```
## Question 1.1.3
It is surprising that the pclass has no missing values. Because if people didn't survive, how can you know which class is he or she in? Therefore, I find it surprising not finding any NA values in pclass.

#Question 1.2
## Question 1.2.1
I think the factors that may affect survive is sex, pclass and age.
Sex: i think woman should have a higher chance of survival because people tend to let woman take the rescue boat.
Pclass: I think children should have a higher chance of survival because people tend to let children take the rescue boat.
Age: I think young people should have a higher chance of survival because young people have better health condition to wait for the rescue boat.

## Question 1.2.2
```{r}
titanic <- titanic %>%
  mutate(child = ifelse(age < 14, 1, 0))
```

## Question 1.2.3
There are three classes in pclass variable, including 1, 2, 3. But it does not have a numerical meaning so we do not calculate it. Instead, we can consider it as three categories, first class, second class and third class. Therefore, it is better to make pclass categorical.
```{r}
titanic$pclass <- factor(titanic$pclass)
```

## Question 1.2.4
We can interpret the results by AME
child 18 pct points more likely to survive
pclass2 17 pct points less likely to survive(compared do “pclass1”)
pclass3 31 pct points less likely to survive(compared do “pclass1”)
Male 49 pct points less likely to survive
```{r}
summary(m <- glm(survived ~ pclass+sex+child, data=titanic, family=binomial()))
me <- margins(m)
summary(me)
```
## Question 1.2.5
Woman have larger chances for survival
Young have larger chances for survival
while class1 have larger chances for survival than class2, and class2 have larger chances for survival than class3.
Gender is the factor that affects the survival the most. 


## Question 1.2.6
Create a young man variable
young man are 5.8 pct points less likely to survive
While male 49 pct points less likely to survive, we can find that young man's chances to survive is larger than all man.
```{r}
titanic <- titanic %>%
  mutate(young_man = ifelse(18<age & age<35, 1, 0))
summary(o <- glm(survived ~ young_man, data=titanic, family=binomial()))
summary(margins(o))
```

## Question 1.2.7
Are the survivors’ accounts broadly accurate? Yes, the results matches the explanation in the passage.
Women and children were the first to get to the boats, so women and children have higher survival chances. 
Third class passengers could hardly reach to the boat deck so they have the lowest survival rate of the three classes.

Did the order break down? no it did not break down, woman and child still have higher chances to survive.

Can you tell anything else interesting?
I do find it interesting because in the movie Titanic, Rose, who is a woman in the first class did survive eventually while Jack, who is a man in the third class didn't. The results matched the movie.

# Question 2
## Question 2.1
load the data, find the nas in each column, and only select price, bedrooms,room_type, accommodates
```{r}
airbnb <- read.csv("airbnb-vancouver-bc-listings.csv.bz2")
airbnb1<-airbnb %>%
  select(price, bedrooms,room_type, accommodates)
head(colSums(is.na(airbnb)))
```

## Question 2.2
convert the price into numeric and replace the NAs in bedrooms with 0 bedrooms.
```{r}
airbnb1<-airbnb %>%
  select(price, bedrooms,room_type, accommodates)%>%
  mutate(
    price = as.numeric(str_replace_all(price, "[^0-9.]", "")), 
    bedrooms = ifelse(is.na(bedrooms), 0, bedrooms))
```

## Question 2.3
As the chart shows below, there is a well-defined lower bound but no obvious upper bound. The corresponding distributions tend not to look normal but are more similar to log-normal. I think it is better to do a log transformation to make it more similar to normal. Also, log transformation improves the predictive power of the model (increases R2)
```{r}
ggplot(airbnb1, aes(x = price)) +
  geom_histogram(binwidth = 100, fill = "blue", color = "white") +
  labs(title = "Distribution of Airbnb Prices", x = "Price", y = "Frequency") +
  theme_minimal()

ggplot(airbnb1, aes(x = log(price))) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "white") +
  labs(title = "Distribution of Log-Transformed Airbnb Prices", x = "Log(Price)", y = "Frequency") +
  theme_minimal()
```

## Question 2.4
Add new column named bedrooms_category
```{r}
airbnb1 <- airbnb1 %>%
  mutate(
    bedrooms_category = cut(
      bedrooms,
      breaks = c(0, 1, 2, 3, Inf),  
      labels = c("0", "1", "2", "3+"),
      include.lowest = TRUE,
      right = FALSE
    )
  )
```

## Question 2.5
model1 using log(price) R2=30%
model using price R2=8%
"model1" using log(price) has higher R2 than "model" using price
In model2, the R2 is 43%, room_type_entire room is the reference category
```{r}
model<- lm(price ~ bedrooms_category, data = airbnb1)
summary(model)
model1 <- lm(log(price) ~ bedrooms_category, data = airbnb1)
summary(model1)

model2 <- lm(log(price) ~ bedrooms_category+room_type+accommodates, data = airbnb1)
summary(model2)

```

## Question 2.6
accommodates means how many people can live in the airbnb, we can see that most airbnb can accommodate 2 people
room type here can be categorized in to four categories, Entire home/apt, Hotel room, Private room and shared room. Most of the airbnb are entire home. 
```{r}
table(airbnb1$accommodates)
table(airbnb1$room_type)
```

## Question 2.7
accommodates_category converts accommodates into three categories "1", "2" and "3 or more"
room_type_category converts room_types into three categories "Entire home/apt", "Private room" and "Other"
```{r}
airbnb1 <- airbnb1 %>%
  mutate(
    accommodates_category = cut(
      accommodates,
      breaks = c(1, 2, 3, Inf),  
      labels = c("1", "2", "3 or more"),
      include.lowest = TRUE,
      right = FALSE))%>%
  mutate(
    room_type_category = case_when(
      room_type == "Entire home/apt" ~ "Entire home/apt",
      room_type == "Private room" ~ "Private room",
      TRUE ~ "Other"
    )
  )
```

## Question 2.8
room_type_entire room is the reference category
R2 here is 40%
In model3, instead of bedrooms_category2 and room_type_category1. Other variables are significant, the p-values are small which means, the H0 is unlikely to happen, and we could reject the null hypothesis. 

when bedrooms_category2 increases by one unit, the log(price) will increase by 4.4+0.17

when bedrooms_category3+ increases by one unit, the log(price) will increase by 4.4+0.38
when room_typePrivate room increases by one unit, the log(price) will decrease by 4.4-0.43

when room_typeShared room increases by one unit, the log(price) will decrease by 4.4-0.45

when accommodates_category2 increases by one unit, the log(price) will increase by 4.4+0.3

when accommodates_category3 increases by one unit, the log(price) will increase by 4.4+0.46
```{r}
model3 <- lm(log(price) ~ bedrooms_category+room_type_category+accommodates_category, data = airbnb1)
summary(model3)
```

## Question 2.9
Here the room type category "other" is the combination of hotel and shared rooms.
The P-value here is 0.415,which means the H0 may have happen, so we cannot reject the null hypothesis. That is to say, there isn't a relationship between the room type other and the price. Therefore, room_type_categoryOther it is not statistical significant. 


## Question 2.10
use model3 to predict the (log) price for each listing
```{r}
predicted_log_price<-(predict(model3, newdata = airbnb1))
head(predicted_log_price)
```

## Question 2.11
to calculate the residual, we have to use the actual log(price)-the predicted_log_price
then we can calcualte the rmse using the formula: sqrt(mean(residuals^2))
```{r}
residuals <- log(airbnb1$price) - predicted_log_price
rmse <- sqrt(mean(residuals^2))
rmse
```

## Question 2.12
filter bedrooms=2 and accommodates=4 to get the predicted log price
```{r}
data_12<-airbnb1%>%
  filter(bedrooms==2, accommodates==4)
predicted_log_price_12<- predict(model3, newdata = data_12)
head(predicted_log_price_12)
```

I spend around 10 hours on this assignment